### 2025-01-29 06related_conclusion.tex
.

\section{Related Work}
Categorical quantum mechanics (CQM)~\cite{DBLP:conf/lics/AbramskyC04}
is a category-theoretic foundation for quantum mechanics. CQM is the
basis of the ZX calculus~\cite{DBLP:conf/icalp/CoeckeD08}, a popular
graphical calculus for quantum states~\cite{vandewetering2020zx}. A
key aspect of the ZX calculus and related graphical calculi is that
they can be equipped with powerful rewriting systems. These systems
support a form of graph rewriting, i.e.\, they exploit the graph
representation of quantum states. There exists several completeness
results for ZX-like rewriting systems, as well as some practical
tools, such as PyZX~\cite{kissinger2019pyzx}. Similarly, there exist
completeness results and tools for the sum-over-paths representation
of quantum states~\cite{amy2018towards,amy2023complete}.

\paragraph*{Comparison with DiracDec~\cite{diracdec}}
Xu \emph{et al}~\cite{diracdec} define a language and an associate and
commutative rewriting system for Dirac notation, and implement their
language and rewriting system in Mathematica. Our approach follows a
similar pattern. However, there are significant differences in terms
of scope, expressiveness, and efficiency. We review these differences
below.

The most obvious difference is that D-Hammer targets labelled Dirac
notation, whereas DiracDec targets plain Dirac notation. As already
mentioned, the use of labelled Dirac notation is essential in many
applications; in particular, labelled Dirac notation can simplify
notation and proofs, and is in general better suited for writing and
reasoning about complex, many-body, quantum states. However, there are
several other important differences. First, our language leverages the
power of higher-order dependent type theory. In particular, we
leverage higher-order functions to provide a compact and expressive
representation of big operators. Second, our language supports AC
symbols with indefinite arities, which leads to compact representation
of terms, and eases AC reasoning. Third, the relation with Mathematica
is fundamentally different. DiracDec is implemented in Mathematica; as
a consequence, its behavior, and in particular, its typing rules are
constrained by the type system of Mathematica. In contrast, D-Hammer
is designed as a separate tool; as a consequence, it benefits from an
improved representation of terms (e.g.\, AX operators with indefinite
arity), a more expressive type system (e.g.\, dependent types), and a
more efficient rewriting engine. D-Hammer still relies on Mathematica
to reason about functions that are not natively supported by rewriting
rules, but these interactions are constrained and do not have negative
effects on the overall efficiency of the system. This is reflected in
our experimental comparison of D-Hammer and DiracDec, which shows how
the former outperforms the latter.

%% The first step is to formalize the language for Dirac notation. In DiracDec, the language has a concrete design, where the same syntax for different types corresponds to distinct symbols. Our goal is to transition from this concrete design to a more abstract one, which aligns more closely with the conventional Dirac notation we use and simplifies the term rewriting system.

%% %% Automated theorem proving has seen significant advancements in recent years, particularly through the development of satisfiability modulo theories (SMT) solvers,
%% %% % SMT solvers extend propositional satisfiability by integrating theories such as arithmetic, arrays, and bit-vectors, making them suitable for solving a wide range of verification problems in both hardware and software domains. 
%% %% including prominent tools like Z3. These solvers have become essential in various fields such as formal verification, synthesis, and model checking. 
%% %% Equational reasoning is another crucial area of research within automated theorem proving, which focuses on solving problems that involve equations between terms in an algebraic structure. 
%% %% Equational provers, such as Vampire and E, have played a pivotal role in addressing the challenge of proving equations in first-order logic, employing sophisticated algorithms like superposition and term rewriting.

%% %% Formal verification of quantum computation is receiving increasing attentions during these years. See~\cite{Lewis2023} for a comprehensive review. 
%% %% Verification frameworks in Coq include the foundational formalization CoqQ and quantum circuit language QWIRE. 
%% %% Verifications of quantum programs are also considered, such as the Hoare logic based methods [] and model checking based methods [].
%% %% The equational reasoning of Dirac notation is crucial for establishing property proofs in these works.
%% %% Verification theoreis and tools based on other languages are also proposed, such as PyZX for reasoning and simplification of ZX-calculus.

\section{Conclusion and Future Work}
We have designed and implemented D-Hammer, a dependently typed
higher-order language and proof system for labelled Dirac
notation. D-Hammer benefits from an optimized implementation in
\CC\ and a tight integration with Mathematica to reason about a broad
range of mathematical functions, including trigonometric and
exponential functions, that are commonly used in quantum physics.
There are three important directions for future work. The first
direction is to enhance D-Hammer with support for big tensor notation,
i.e.\, to support expressions of the form $\bigotimes_{i\in I}
e_i$. The second direction is to extend D-Hammer with a mechanism to
generate independently verifiable certificates.  There is a large body
of work on producing certificates for automated tools, in particular
SMT solvers; see e.g.~\cite{DBLP:journals/cacm/BarbosaBCDKLNNOPRTZ23}
for a recent overview. One potential option would be to integrate
D-Hammer with the Coq or Lean proof assistants; in the first case, one
would benefit from the formalization of labelled Dirac notation in
CoqQ~\cite{Zhou2023}, whereas in the second case, one would benefit
from powerful mechanisms to integrate rewriting procedures into the
Lean proof assistant. Finally, the third direction is to connect
D-Hammer with quantum program verifiers. Two potential applications
are automating equational proofs for tools that already use Dirac
notation, and to substitute numerical methods for tools that use
matrices instead of symbolic assertions.

%% Based on the first equational reasoning tool for Dirac notation called DiracDec, this work improves and extends the theory for practical applications, and provides the solver D-Hammer. Experiments show that the tool demonstrates advantages in decidability, efficiency and usability.

%% We expect D-Hammer to have applications in areas like quantum program verification or proofs of post-quantum cryptography protocols in the future.
%% One promising following up is to connect D-Hammer with theorem provers like Coq. It involves transforming theorem prover expressions into D-Hammer, and verify the proof trace of D-Hammer in theorem provers. Besids, most quantum program verifiers nowadays depend on matrix calculations. D-Hammer can serve as the replacement for matrix methods to enable symbolic deductions.