### 2025-01-30 05tool_eval.tex
update evaluation

\section{Implementation and Case Study}
The refinements and extensions above concludes in our implementation called D-Hammer, a solver written in \CC. It has a parser built by ANTLR4, and scalar reasonings are powered by a Mathematica kernel. The user can use commands to make definitions and assumptions in the maintained context, conduct the normalization and equivalence checking, and obtain the rewriting trace output. 
It can be used from the command line interactively, or can be integrated into other \CC\ projects as a library.
This implementation is tested on the benchmark of the previous work and new examples for labelled Dirac notation, demonstrating significant improvement in expressivity and efficiency.

\subsubsection{Project Structure}
The project structure is illustrated in~\Cref{fig: dhammer structure}.
\texttt{ualg} is the module for universal algebra, defining basic concepts like terms and substitutions. It serves as the library for \texttt{dhammer}, which are then utilized in the example benchmarks and the toplevel command line application. The components of \texttt{dhammer} are as follows:
\begin{itemize}
    \item \texttt{symbols.cpp}: the reserved symbols and AC symbols;
    \item \texttt{syntax\_theory.cpp}: syntax related algorithms, such as de Bruijn normalization and freeness of variables;
    \item \texttt{calculus.cpp}: type checker and intergration with Mathematica;
    \item \texttt{reduction.cpp}: all the rewriting rules and transformations;
    \item \texttt{dirac\_parser.cpp}: parser for Dirac notation and D-Hammer commands;
    \item \texttt{prover.cpp}: the prover that maintains the context and process commands like definition or equivalence checking.
\end{itemize}

\input{fig/proj_structure.tex}

The internal data structure for terms is a pointer-based syntax tree following the function application style:
\[
    \texttt{
        term ::= ID | ID [term (, term)*].
    }
\]
The syntax tree can be an identifier, or an application with an identifier as the function head, and several syntax trees as arguments. There are several Dirac notation terms and their corresponding syntax trees.
\footnotesize{
\begin{align*}
    & X_1 + X_2 + X_3 && \texttt{ADD[X1, X2, X3]} 
    \\
    & \lambda x: \OType(T_1,  T_2). x^\dagger && \texttt{FUN[x, OTYPE[T1, T2], ADJ[x]]}
    \\
    & \sum_{i \in \mathbf{U}(T)} \ket{i} \bra{i} && \texttt{SUM[USET[T], FUN[i, BASIS[T], OUTER[KET[i], BRA[i]]]]}
\end{align*}
}
The syntax tree structure is also compatible with the datatype of Mathematica. This improves the interoperability between D-Hammer and the Mathematica system, enabling them to work interleavingly.
To improve usability, D-Hammer also supports many special notations for terms,  and most Dirac notation terms will be encoded in the natural way.
Here are some examples for the parsing syntax.

\begin{figure}
    \center
\begin{tabular}{c >{\centering\arraybackslash}p{4cm} l}
    \hline
    syntax & parsing result & explanation \\
    \hline
    \texttt{|e>} & \texttt{KET[e]} & the ket basis\\
    \texttt{e1 + ... + en} & \texttt{ADD[e1, ..., en]} & the addition\\
    \texttt{e1\ e2} & \texttt{COMPO[e1, e2]} & composition in Dirac notation \\
    \texttt{e1\^{}*} & \texttt{CONJ[e1]} & scalar conjugation \\
    \texttt{fun i : T => X} & \texttt{FUN[i, T, X]} & lambda abstraction \\
    \hline
\end{tabular}
\end{figure}

Finally, D-Hammer uses a prover to host the computation. The prover maintains a well-formed context $E$, and processes commands to modify the context and conduct calculations. The commands are listed below.
\begin{itemize}
    \item \texttt{\textcolor{NavyBlue}{Def} ID := term.} It defines the \texttt{ID} as the \texttt{term}, using the \textbf{W-Def} typing rule.
    \item \texttt{\textcolor{NavyBlue}{Var} ID := term.} It make an assumption of \texttt{ID} with the \texttt{term} as type, using the \textbf{W-Assume} typing rules.
    \item \texttt{\textcolor{NavyBlue}{Check} term.} Type checking the \texttt{term} and output the result.
    \item \texttt{\textcolor{NavyBlue}{Normalize} term.} Normalize the \texttt{term} using the algorithm introduced in~\Cref{sec: decide}.
    \item \texttt{\textcolor{NavyBlue}{CheckEq} term \textcolor{NavyBlue}{with} term.} Check the equivalence of the two terms calculating and comparing their normal forms.
\end{itemize}
The prover will type check the terms for each command. We can also use \texttt{\textcolor{NavyBlue}{Normalize} term \textcolor{NavyBlue}{with trace}.} to output the proof trace during normalization. The proof trace is a sequence of records, including the rule or transformation appied, the position of application, and the pre- and post-transformation terms. The record helps understand the normalization procedure better, and can be turned into verified proofs in theorem provers in the future.

\subsection{Equivalence Checking Algorithm}

The equivalence of Dirac notations are decided by normalization, which transforms equivalent expressions under a set of axioms into the same syntax. We use an efficient algorithm to perform the normalization totally on $R \cup F$.

\begin{enumerate}
    \item \textbf{First Rewritings}: Expand definitions and simplify expressions.
    \item \textbf{Variable Expansion}: Convert to abstract element-wise representation.
    \item \textbf{Second Rewritings}: Normalize terms on \( R \) modulo \( F \).
    \item \textbf{Sorting Without Bound Variables}: Normalize AC-equivalence.
    \item \textbf{Swap Successive Summations}: Normalize SUM-SWAP equivalence.
    \item \textbf{De Bruijn Normalization}: Normalize \( \alpha \)-equivalence.
\end{enumerate}
Here steps 1 - 3 are term rewriting for $R$, and steps 4 - 6 are special algorithms to further normalize the axiom set $F$. Step 2 expands variables like $K$ to their abstract element-wise representation $\sum_{i} (\bra{i} \cdot K). \ket{i}$, which improves the decidability with sum. For step 4 and 5, the key observation is that in a successive sum expression \( \sum_{i \in s_1} \cdots \sum_{j \in s_n} A \), the names and order of the bound variables \( i, \dots, j \) can be freely permuted. Therefore we first ignore bound variables and normalize AC-equivalence by sorting. Then decide the order of summation accordingly. The last step converts functions to de Bruijn indeices~\cite{deBruijn1972lambda} to decide $\alpha$-equivalence.

As a comparison, the previous work conducts normalization paritally on $R$, and decide axioms in $F$ by iterating through all possible permutations.
Our algorithm totally normalize the term and significantly improves the efficiency.
\[
\text{(partial)}\qquad e_1 \mathop{\twoheadrightarrow}^R e_1' \mathop{=}^F e_2' \mathop{\twoheadleftarrow}^R e_2
\qquad
\text{(total)}\qquad e_1 \mathop{\twoheadrightarrow}^{R \cup F} e \equiv e \mathop{\twoheadleftarrow}^{R \cup F} e_2
\]

\yx{we can add a normalization example here.}

\subsection{Use Case}
Here we encode the motivating~\Cref{ex: motivating}, examine and explain how to check it using D-Hammer. The encoding is shown below.

    \begin{lstlisting}[style=dhammer]
Var T : INDEX. Var M : OTYPE[T, T].
Def phi := idx T => Sum nv in USET[T], |(nv, nv)>.
Var r1 : REG[T]. Var r2 : REG[T].
CheckEq M_r1 (phi T)_(r1, r2) with (TPO T T M)_r2 (phi T)_(r1, r2).
    \end{lstlisting}

The first three lines use the \texttt{\textcolor{NavyBlue}{Var}} and \texttt{\textcolor{NavyBlue}{Def}} commands to set up the context for the Dirac notation.
\texttt{T} is a type index, representing arbitrary Hilbert space types. \texttt{M} is assume to be an operator in the Hilbert space with type \texttt{T}. \texttt{phi} is defined as the maximally entangled state, depending on the bound variable \texttt{T} as index.
\texttt{r1} and \texttt{r2} are register names for the two subsystems.

In the left hand side of \texttt{\textcolor{NavyBlue}{CheckEq}} command, \texttt{M\_r1} denotes the labelled notation $M_{r_1}$, and \texttt{(phi T)\_(r1, r2)} denotes the entangled state $\ket{\Phi}_{(r_1, r_2)}$. They are connected by a white space, which is parsed into the composition of Dirac notation, and will be reduced into the operator-ket multiplication after typing. The right hand side is interpreted similarly, except the defined symbol \texttt{TPO} in the environment:

\begin{lstlisting}[style=dhammer]
Def TPO := idx sigma => idx tau => fun O : OTYPE[sigma, tau] => Sum i in USET[sigma], Sum j in USET[tau], (<i| O |j>).(|j> <i|).
\end{lstlisting}

The \texttt{TPO} symbol represents the transpose of operators, and encodes the formalization in~\Cref{ex: formalizing motivating}. Thanks to the design of environment and functions, many other commonly used symbols in Dirac notation are encoded and provided as defined symbols in D-Hammer.

Within one second, the prover reports the result of equivalence with their common normal form:
    \begin{lstlisting}[style=dhammer]
The two terms are equal.
[Normalized Term] SUM[USET[T], FUN[BASIS[T], SUM[USET[T], FUN[BASIS[T], SCR[DOT[BRA[$\texttt{\$1}$], MULK[M, KET[$\texttt{\$0}$]]], LTSR[LKET[$\texttt{\$1}$, r1], LKET[$\texttt{\$0}$, r2]]]]]]] : DTYPE[RSET[r1, r2], RSET]
    \end{lstlisting}

The normal form is in the internal syntax tree format mentioned above. A more readable interpretation is:
\[
\sum_{\mathbf{U}(T)} \sum_{\mathbf{U}(T)} \bra{\$1}M\ket{\$ 0} . \ket{\$1}_{r_1} \otimes \ket{\$0}_{r_2} : \DType(\{r_1, r_2\}, \emptyset).
\]
Here $\$0$ and $\$1$ are de Bruijn indices. The result is a ket on the $\{r_1, r_2\}$ system as expected, and follows pattern proposed in~\Cref{sec: labelled}.

\section{Evaluation}
We evaluate D-Hammer on several example sets, and make a comparison with the previous tool DiracDec~\cite{diracdec}.
The experiments are carried out using a MacBook Pro with M3 Max chip. Results are summarized as follows.

{
    \center
    \begin{tabular}{c|c c c|c c c}
        \hline
        \multirow{2}{*}{source} & \multicolumn{3}{c|}{DiracDec} & \multicolumn{3}{c}{D-Hammer} \\
        \cline{2-7}
                                 & expressable & success & time(s)           & expressable & success & time(s)                 \\
        \hline
        textbook(QCQI)          & 18          & 18        &    1.02        &    18      & 18          &   0.82      \\
        CoqQ                    & 162          & 156       &    48.69       &   158     &  158   &     9.74     \\
        % others                  & 7          &  6         &   77.20    &    7        &   6     &  2.13     \\
        % circuits                 & 2          & 2       &    17.67       &   3     &  2   &     1.4     \\
        % research paper                & 4          & 4         &  59.53       &   4    & 4       &  0.73     \\
        labelled Dirac notation      &   -         &   -          &     -           &      -      &        18       &     22.72    \\
        \hline
    \end{tabular}        
}

\subsubsection{Textbook (QCQI)}
The 18 examples of QCQI textbook equations from DiracDec are all encoded and solved. The running time are short and close because these examples are simple and direct.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{fig/coqq.pdf}
    \caption{Time comparison between DiracDec and D-Hammer on the CoqQ benchmark.}
    \label{fig: CoqQ plot}
\end{figure}

\subsubsection{CoqQ}
As to expressibility, D-Hammer encodes 158 examples for the CoqQ part, 4 less than DiracDec. This is because the language for DiracDec has the support for projectors $\texttt{fst}$ and $\texttt{snd}$ on basis pairs, satisfying $\texttt{fst} (s, t) = s$ and $\texttt{snd} (s, t) = t$. We found this feature is almost not used and removed the support. For decidability, we improved the rewriting rules about sum for D-Hammer, and proved several more examples failed by DiracDec. 
\Cref{fig: CoqQ plot} shows a direct comparison of their time efficiency.
We can observe that for small examples, D-Hammer is slower because of the marginal overhead. 
When the running time goes up, D-Hammer provides 2 to 40 times of acceleration in average.
The relationship between efficiency improvement and example structure further supports the necessity of our algorithm. 
We find that examples with great efficiency improvement, e.g. COQQ-129 and COQQ-148 shown in~\Cref{fig: CoqQ plot}, tend to use deeply nested summations, where our algorithm has an effect.

% \subsubsection{Others}
% Other examples are taken from quantum circuits and one recent paper [TODO]. Because these examples involve decomposition on concrete $\ket{0}$ and $\ket{1}$ qubit basis, resulting a lot of addition elements, our algorithm dealing with AC symbols brings along significant improvement.
% One typical example is
% \begin{gather*}
%     (I \otimes P) \cdot U \cdot (I \otimes P) \cdot U^\dagger = \ket{0}\bra{0} \otimes I + \ket{1}\bra{1}\otimes (P \cdot P), \\
%     \textrm{where } P \triangleq e^{-i\theta/2} \ket{0}\bra{0} + e^{i\theta/2}\ket{1}\bra{1}, \qquad U \triangleq \ket{0}\bra{0} \otimes X + \ket{1}\bra{1} \otimes I.
% \end{gather*}
% It takes DiracDec about one minute, but D-Hammer solves it within one second.

\subsubsection{Labelled Dirac Notation}
We present a new set of examples for labelled Dirac notation (LDN), as detailed in~\Cref{sec: examples for labelled}. These examples, illustrated in~\Cref{fig: labelled examples}, include six representative cases drawn from a variety of sources, such as well-established theorems, research paper results, quantum circuit equivalence, and quantum program verification.
D-Hammer normalizes them and successfully checks their equivalence using the algorithm in~\Cref{sec: labelled}.
Among the examples, LDN-12 shows the flexibility in combining labelled Dirac notations.
Example LDN-14 and LDN-18 are about circuits or programs on qubits. LDN-14 shows how to calculate controlled-not gate in different ways. In LDN-18, $q, q_1$ and $q_2$ represent variables for quantum programs, and the equation shows the change of the state after applying $X$ gate on $q_2$. Experiments show that our algorithm is relatively inefficient for labelled Dirac notations based on qubits.

It is a noteworthy result that D-Hammer solved LDN-10, a highly complex and lengthy example. It is a theorem on quantum separation logic from~\cite{DBLP:conf/lics/ZhouBHYY21}, and is challenging to prove even for experts. Notice that 7 registers are involved here, and it would be impractical to organize and referring to the subsystems without using labels.

\begin{figure}[h]
    \center
    % \scriptsize
    \setlength{\extrarowheight}{2pt}
    \begin{tabular}{l c c l}
        \hline
        example & source & time(s) & equation \\
        \hline
        LDN-4 & theorem & 0.03 & \( \bra{\Phi}_{(x, y)} M_y \cdot N_x \ket{\Phi}_{(x, y)} = \mathrm{tr}(M^T N) \) \\
        LDN-10 & paper~\cite{DBLP:conf/lics/ZhouBHYY21} & 5.17 & \( \mathrm{tr}_{((a',(b,b')),c')}\Big[\mathrm{tr}_{r}\Big(U_{(r,(a,b))} \cdot\Big(|s\>_r\<s|\otimes 
        \Big[V_{((a',(b,b')),c')} \cdots \)\\
        LDN-11 & paper~\cite{PALSBERG2024206} & 0.12 & \( U_{(a,b)}\cdot W_{(b,c)}\cdot V_{(a,c)} = 
        \sum_i |i\>_a\<i|\otimes \big((P_i)_c\cdot W_{(b,c)}\cdot (Q_i)_c\big)\) \\
        LDN-12 & circuit & 0.07 & \( \ket{i}_{a;b}\bra{j} \cdot C_{(b,c)} \cdot D_{(c,d)} = {}_b\bra{j} \cdot C_{(b,c)} \cdot D_{(c,d)} \cdot \ket{i}_{a}\) \\
        LDN-14 & circuit & 7.37 & \scriptsize{\( \textsf{CNOT}_{rq}\ket{\textrm{GHZ}}_{pqr} = (\textsf{CNOT}\ket{00})_{rq}\ket{0}_p + (\textsf{CNOT}\ket{11})_{rq}\ket{1}_p \)} \\
        LDN-18 & program & 1.01 & \( -\ket{0}_q\ket{+-}_{q_1, q_2} = X_{q_2} \ket{0}_q\ket{+-}_{q_1, q_2} \) \\
        \hline
    \end{tabular}
    \caption{Part of examples for labelled Dirac notations.}
    \label{fig: labelled examples}
\end{figure}